---
title: "Data_input_preproccess_for_scPagwas"
date: "Last Updated: `r format(Sys.time(), '%d, %B, %Y at %H:%M')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_input_preproccess_for_scPagwas}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
  fig.path = "figures/vignette-",
 out.width = "60%"
)
```

# Data input

There were three input resources for scPagwas: scRNA-seq dataset(seruat format), a GWAS summary dataset(txt file) on a given phenotype and an extensive panel of pathways or functional gene sets(gene symbol list).

## 1.Single cell data Input

The healthy BMMC scRNA-seq dataset was downloaded from the website: <https://jeffgranja.s3.amazonaws.com/MPAL-10x/Supplementary_Data/Healthy-Data/scRNA-Healthy-Hematopoiesis-191120.rds>

-   1). The scRNA-seq dataset should have Idents for cell types or clusters.
-   2). The scRNA-seq dataset should be normalized.

### 1.1.read the scRNA-seq dataset(seruat format)

```{r message=FALSE, eval = FALSE}
library("scPagwas")
library("Seurat")
library("scRNAseq")
library("SingleCellExperiment")
library("stringr") 
scRNA_Healthy_Hema<-readRDS("E:/OneDrive/SingleCell/data/PBMCscATAC-seq/scRNA-Healthy-Hematopoiesis-191120.rds")
```

Because the Example data `scRNA_Healthy_Hema` is SingleCellExperiment format, here creating a seruat format data:

```{r message=FALSE, eval = FALSE}
counts <- assay(scRNA_Healthy_Hema, "counts")
Seu_Healthy_Hema <- CreateSeuratObject(
               counts = counts, 
               meta.data=as.data.frame(colData(scRNA_Healthy_Hema)),
               min.cells = 3, 
               min.features = 200)
```

### 1.2.Set Idents

```{r message=FALSE, eval = FALSE}
Idents(Seu_Healthy_Hema)<-scRNA_Healthy_Hema@colData$BioClassification
table(Idents(Seu_Healthy_Hema))

        01_HSC 02_Early.Eryth  03_Late.Eryth  04_Early.Baso    05_CMP.LMPP       06_CLP.1 
          1425           1653            446            111           2260            903 
        07_GMP    08_GMP.Neut         09_pDC         10_cDC 11_CD14.Mono.1 12_CD14.Mono.2 
          2097           1050            544            325           1800           4222 
  13_CD16.Mono         14_Unk       15_CLP.2       16_Pre.B           17_B      18_Plasma 
           292            520            377            710           1711             62 
      19_CD8.N      20_CD4.N1      21_CD4.N2       22_CD4.M      23_CD8.EM      24_CD8.CM 
          1521           2470           2364           3539            796           2080 
         25_NK         26_Unk 
          2143            161 
```

### 1.3.Normalized

Scale and Normalize the dataset.

```{r setup, eval = FALSE}
Seu_Healthy_Hema <- ScaleData(Seu_Healthy_Hema)
Seu_Healthy_Hema <- NormalizeData(Seu_Healthy_Hema, normalization.method = "LogNormalize", scale.factor = 10000)
```

## 2.GWAS summary data Input

GWAS Summary statistics are download from: <ftp://ftp.sanger.ac.uk/pub/project/humgen/summary_statistics/UKBB_blood_cell_traits/> for UK Biobank.

-   basophil cell count: ieu-b-29
-   eosinophil cell count : ieu-b-33
-   lymphocyte cell count: ebi-a-GCST004627
-   monocyte cell count: ieu-b-31
-   neutrophil cell count: ieu-b-34
-   white blood cell count: ieu-b-30
-   Lymphocyte percentage of white cells: ebi-a-GCST004632
-   Hemoglobin concentration: ebi-a-GCST90002311
-   Mean corpuscular hemoglobin concentration: ebi-a-GCST90002329
-   Mean corpuscular volume: ebi-a-GCST90002335

The GWAS Summary statistics file need to be processed into a "txt" file including six coloumn and tab-delimited.

### 2.1.Check the example GWAS Summary statistics file

```{r message=FALSE}
library(scPagwas)
gwas_data <- bigreadr::fread2(system.file("extdata", "GWAS_summ_example.txt", package = "scPagwas"))
knitr::kable(head(gwas_data))
```

### 2.2.Take the monocyte count as real example and deal with it step by step:

Preprogress the raw monocytecount GWAS Summary statistics file.
Extract the useful information from raw files.
Download the Raw gwas GWAS Summary statistics file: <https://gwas.mrcieu.ac.uk/files/ieu-b-31/ieu-b-31.vcf.gz>

```{r message=FALSE, eval = FALSE}
library(readr)
############read the raw vcf file
monocytecount<-"ieu-b-31.vcf.gz"
GWAS_raw <-read_table2(monocytecount,comment = "#")
colnames(GWAS_raw )<-c("CHROM","POS","ID","REF","ALT","QUAL","FILTER","INFO","FORMAT","IEU")
#ES:SE:LP:AF:SS:ID
#ieu-a type of file:split the FORMAT and IEU coloumn, which including different information.
value_list<-lapply(1:nrow(GWAS_raw),function(i){
  index <- unlist(strsplit(GWAS_raw$FORMAT[i], split = ":",fixed=T))
  value <- unlist(strsplit(GWAS_raw$IEU[i], split = ":",fixed=T))
  names(value)<-index
  value<-value[c("ES","SE","LP","AF","SS")]
  return(value)
  })

 value_df<- matrix(as.numeric(unlist(value_list)),byrow=T,ncol=5)
 value_df<-as.data.frame(value_df)
 colnames(value_df)<-c("ES","SE","LP","AF","SS")
 gwas_data<-GWAS_raw[,c(1:5)]
 gwas_data<- cbind(gwas_data,value_df)
 colnames(gwas_data)
#[1] "CHROM" "POS"   "ID"    "REF"   "ALT"   "ES"    "SE"    "LP"    "AF"   
#[10] "SS"
#change the colnames:
colnames(gwas_data)<-c("chrom","pos","rsid","REF","ALT","beta","se","maf","N")
#select the useful information for scPagwas.
gwas_data<-gwas_data[,c("chrom","pos","rsid","beta","se","maf")]
write.table(gwas_data,file="monocytecount_gwas_data.txt",row.names=F,quote=F)
```

## 3.Pathway gene list

There are some processed pathway gene list provided by scPagwas package: all these pathway list are including gene symbols.

-   Genes_by_pathway_kegg
-   genes.by.celltype.pathway
-   genes.by.gpbp.pathway
-   genes.by.reactome.pathway
-   genes.by.regulatory.pathway
-   genes.by.tft.pathway
-   genes.by.hallmark.pathway
-   genes.by.immunologic.pathway
-   genes.by.immunesigdb.pathway

```{r message=FALSE}
Genes_by_pathway_kegg[1:3]
```

There are some steps for get these pathway gene list: There is no need to repeat running these code.


### KEGG pathway list

```{r message=FALSE, eval = FALSE}
library(KEGGREST)
pathways.list <- keggList("pathway", "hsa")
# Pull all genes for each pathway
pathway.codes <- sub("path:", "", names(pathways.list))	
genes.by.pathway_kegg <- sapply(pathways.list,
                           function(pwid){
                             pw <- keggGet(pwid)
                             if (is.null(pw[[1]]$GENE)) return(NA)
                             pw2 <- pw[[1]]$GENE[c(FALSE,TRUE)] # may need to modify this to c(FALSE, TRUE) for other organisms
                             pw2 <- unlist(lapply(strsplit(pw2, split = ";", fixed = T), function(x)x[1]))
                             return(pw2)
                           })
```

### Other pathway list

These pathways data are download from [GSEA](https://www.gsea-msigdb.org/gsea/index.jsp) website.

```{r message=FALSE, eval = FALSE}
x <- readLines("c8.all.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.celltype.pathway <- lapply(res, "[", -c(1:2))

x <- readLines("c3.tft.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.tft.pathway <- lapply(res, "[", -c(1:2))

x <- readLines("c5.go.bp.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.gpbp.pathway <- lapply(res, "[", -c(1:2))

x <- readLines("c3.all.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.regulatory.pathway <- lapply(res, "[", -c(1:2))

x <- readLines("h.all.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.hallmark.pathway <- lapply(res, "[", -c(1:2))

x <- readLines("c2.cp.reactome.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.reactome.pathway <- lapply(res, "[", -c(1:2))
save(genes.by.reactome.pathway,file = "genes.by.reactome.pathway.RData")

#all immunologic signature gene sets as Gene Symbols
x <- readLines("c7.all.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.immunologic.pathway <- lapply(res, "[", -c(1:2))

#ImmuneSigDB gene sets as Gene Symbols
x <- readLines("c7.immunesigdb.v7.5.1.symbols.gmt")
res <- strsplit(x, "\t")
names(res) <- vapply(res, function(y) y[1], character(1))
genes.by.immunesigdb.pathway <- lapply(res, "[", -c(1:2))

```

Note:

1.  Sometimes, there is no need to change the pathway list frequently once you choose a fitable pathway list.
2.  The summed number of genes in pathway list should not too much to cost a long time or memory.Once your pathway list is too big, you can choose some sub-pathway suitable for you analysisi or remove some redundance pathways.

### Reduce pathway gene list

```{r message=FALSE}
 length(genes.by.reactome.pathway)
 length(unlist(genes.by.reactome.pathway))
```

We set the `genes.by.reactome.pathway` for example, there are 1615 gene list and 89476 genes, a mount of resouces and time will be cost when running scPagwas, we provid `reduce_pathway` to reduce the pathway list:

-   `pathway_seed` choose a list of pathway names as seed, which are some pahtways cann't be remove.
-   `remove_proporion` The propotion of duplicated between seed pathway and the others

```{r message=FALSE, eval = FALSE}
reduce_genes.by.reactome.pathway<-scPagwas::reduce_pathway(pathway_seed=names(genes.by.reactome.pathway)[sample(1:length(genes.by.reactome.pathway),500)],
                                                 pathway_list=genes.by.reactome.pathway,
                                                 remove_proporion=0.6)
length(reduce_genes.by.reactome.pathway)
#[1] 1045
length(unlist(reduce_genes.by.reactome.pathway))
#[1] 34980
```

## 4.Gene block annotation

Gene block annotation in chromosome is need to prepare for scPagwas.
scPagwas can provide a block annotation data for protein-coding genes.

the processed block annotation data are show here:

```{r message=FALSE}
suppressMessages(library(scPagwas))
knitr::kable(head(block_annotation))
```

Here is the procedure of obtaining the data:

```{r message=FALSE, eval = FALSE}
library("rtracklayer")
gtf_df<- rtracklayer::import("gencode.v34.annotation.gtf.gz")
gtf_df <- as.data.frame(gtf)
gtf_df <- gtf_df[,c("seqnames","start","end","type","gene_name")]
gtf_df <- gtf_df[gtf_df$type=="gene",]
block_annotation<-gtf_df[,c(1,2,3,5)]
colnames(block_annotation)<-c("chrom", "start","end","label")
```


## 5.LD data

The 1,000 Genomes Project Phase 3 Panel was applied to calculate the linkage disequilibrium (LD) among SNPs extracted from GWAS summary statistics.
the processed LD data are show here:


```{r message=FALSE}
suppressMessages(library(scPagwas))
class(chrom_ld)
knitr::kable(head(chrom_ld[["chr1"]]))
```


We use `vcftools` and `plink` to deal with the 1,000 Genomes genotypes.
```{bash eval=FALSE}
./vcftools --vcf ./1000genomes_all_genotypes.vcf --plink-tped --out ./1000genomes_all_genotypes
./plink --tfile ./1000genomes_all_genotypes --recode --out ./1000genomes_all_genotypes
./plink --map ./1000genomes_all_genotypes.map --ped ./1000genomes_all_genotypes.ped --allow-no-sex --autosome --r2 --ld-window-kb 1000 --ld-window-r2 0.2 --out ./ld_1000genome
```

R environment to print out the scPagwas-needed data.

```{r message=FALSE, eval = FALSE}
covid_ld<-read.delim("./ld_1000genome.ld")
#remove sex chrome
covid_ld<-covid_ld[!(covid_ld$ %in% 23),]
colnames(covid_ld)[7]<-"R"
#print out the result in chrom number
lapply(unique(covid_ld$CHR_A), function(i){
  a<-data.table(covid_ld[covid_ld$CHR_A == i,])
  file_name <- paste0("./LD/",i,".Rds")
  saveRDS(a, file = file_name)
})
#integrate the data
chrom_ld<-lapply(as.character(1:22),function(chrom){
  chrom_ld_file_path <- paste(ld_folder, '/', chrom, '.Rds', sep = '')
 ld_data <- readRDS(chrom_ld_file_path)[(R**2 > r2_threshold), .(SNP_A, SNP_B, R)]
  return(ld_data)
})
```
